---
title: 'DS 202 - Lab 4: Tidying and Joining'
subtitle: "Due: Thursday 10/24/2019 by 10 pm"
author: "Matthew Peterson, Shivam Vashi"

date: "10/17/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Outline

- Work in pairs, as in Lab 3
- Create a repository (one per team) & give access to your teammate
- Explore the Cyclone football data. Use `dplyr` to manipulate data frames and `tidyr` to tidy data up. 
- Create an R markdown file that contains all code and your exploratory data analysis
- Submit!

## Github repository

- Owner: Create a GitHub repository called `ds202_lab4`, and give access to the partner.
- For a refreshment, review the procedures for collaboration for Lab 2. 
- Collaborate on `team.Rmd` to finish all the questions.

## Format for submission

- We will check whether or not your lab results are reproducible - i.e. we will knit your `team.Rmd` file. 
- You will also be graded on **style**. Style means is your lab *clear, easy to read, and flows logically*? This lab does **not** need to be in report format but it should be neatly organized with appropriate `R` chunks. 
-  <span style="color:red">Each plot at **minimum** must have labels for both axes and a title. If you don't have these components, you will not get full credit for your answers. Each plot ideally should also come with a description. </span> 

## Submission
Your git repo will contain the following files: `team.Rmd` and `team.html` containing the collaborated work.

- Both teammates need to modify `team.Rmd`. `git log team.Rmd` should show commit information from both teammates.
- To submit: 

  1. Push `team.Rmd` and `team.html` to the owner's git repo.

  2. **Submit both `team.Rmd` and `team.html` to Canvas as well**. In your submission, include: 
    - The owner's GitHub username 
    - The partner's GitHub username 
    - Repo name (should be `ds202_lab4`)
- Due date: Thursday Oct. 24 by 10 pm. 



## Background
An excerpt of the individual player data of the Cyclone football team is available from Canvas. Dowload the [excel file](cyclonesFootball2018.xlsx) and put the file next to your Rmd files. 

Some abbreviations for the column names can be found in the Glossaries sheet in the Excel file.

**References:**

- [Iowa State University Athelete](https://cyclones.com/index.aspx)
- [A Beginner's Guide to American Football](https://www.youtube.com/watch?v=3t6hM5tRlfA) for the rules of American football.

## Loading Data: Cyclone Football 

The Cyclone football data involves multiple Excel sheets and is in a somewhat messy form. The following code loads the Defensive sheet into R. **For this lab, also import the Offensive and the Biography sheets.**

```{r}
library(readxl)
library(tidyverse)
library(tidyr)
library(ggplot2)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
```

```{r}
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
```

## Part one: Cleaning data
The three Excel sheets are loaded into R but they are not useful to us before some cleaning. Manipulate the three imported data frames, so that the cleaned data satisfy the following:

1. The player names and opponent names are factors.
```{r}
defense$Name= factor(defense$Name)
defense$Opponent_Opponent= factor(defense$Opponent_Opponent)

offense$Name= factor(offense$Name)
offense$Opponent_Opponent= factor(offense$Opponent_Opponent)
head(offense)

bio$Name= factor(bio$Name)
head(bio)
```

2. All offensive and defensive statistics are numerical. The `Weight` column in Biography is also numeric. (Hint: look at `?mutate_at` if you want to save writing some code). 
```{r}
head(defense)
bioClean <- bio$Weight <- as.numeric(bio$Weight)

str(defense[2])
defenceClean <- defense
defenceClean$Tackles_Solo <- as.numeric(defense$Tackles_Solo)
defenceClean$Tackles_ASST <-  as.numeric(defense$Tackles_ASST)
defenceClean$Tackles_TFL <-  as.numeric(defense$Tackles_TFL)
defenceClean$Tackles_Sack <-  as.numeric(defense$Tackles_Sack)
defenceClean$Turnover_FF <-  as.numeric(defense$Turnover_FF)
defenceClean$Turnover_FR <-  as.numeric(defense$Turnover_FR)
defenceClean$Turnover_INT <-  as.numeric(defense$Turnover_INT)
defenceClean$Pass_QBH <-  as.numeric(defense$Pass_QBH)
defenceClean$Pass_PB <-  as.numeric(defense$Pass_PB)
head(defenceClean)
```
3. Separate the `Height` column into two columns. Make sure the newly created columns are numeric. Then create one column that represents `Height` (Hint: Use a unit that works well). Drop any irrelavant columns from this process.
```{r,echo=FALSE,results='hide'}
#
separate(bio, "Height", c("Feet", "Inches"), sep = "-" )
```
4. By cleaning the three data frames in R, at the end of this step you should have the following three clean data frames: `defClean`, `offClean`, and `bioClean`. Print the structures of these data frames. **Make sure your data types make sense for each variable.**

## Part two: Tidying (A)
For the following questions, work on the cleaned data frames `defClean`, `offClean`, and `bioClean`.
```{r}
defClean = defense
offClean =offense
bioClean = bio
```
1. Reformat the `defClean` data frame into a tidy format using `gather`. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a **new key column named `Statistic`.** 
```{r}
defClean <- defClean %>% as.data.frame %>% gather(key = Statistic, value = cases, 2:5)
```
2. Create a histogram for each defensive statistic (make sure to explore different binwidths). Use `facet_wrap` to draw all histograms in a single plot. What defensive skills are rare?
```{r}
a <- ggplot(defense, aes(defense$Opponent_Opponent)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Statistic");
b <- ggplot(defense, aes(defense$Tackles_Solo)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust = 1));
c <- ggplot(defense, aes(defense$Tackles_ASST)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust = 1));
d <- ggplot(defense, aes(defense$Tackles_TFL)) + geom_bar()+ theme(axis.text.x = element_text(angle = 90, hjust = 1));
e <- ggplot(defense, aes(defense$Tackles_Sack)) + geom_bar() + theme(axis.text.x = element_text(angle = 90, hjust = 1));
a + facet_wrap(~Tackles_Sack, nrow = 1) + ggtitle("Histograms of defensive statistics")
```
TFL Tackles and sacks are rare to see in the statistics. 

3. Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense. 

```{r, r,echo=FALSE}
Virg = defense[which(defense$Opponent_Opponent == "West Virginia" ),]
Kans = defense[which(defense$Opponent_Opponent == "Kansas State" ),]
print(Virg)
ggplot(Virg, aes(Virg$Name, Virg$Tackles_Solo)) + geom_point() + ggtitle("Tackles by player against West Virginia") + xlab("Player") + ylab("Tackles") + theme(axis.text.x = element_text(angle = 90, hjust = 1));
ggplot(Kans, aes(Kans$Name, Kans$Tackles_Solo)) + geom_point() + ggtitle("Tackles by player gainst Kansas State") + xlab("Player") + ylab("Tackles") + theme(axis.text.x = element_text(angle = 90, hjust = 1));
```
We had a better defense against Kansas state since we had a larger number of solo tackles. 

      (**Hint**: From step 1 we have gathered our `defClean` into 4 columns. \
       Now we only want to focus on solo tackles (hint...`filter`). 
        Look at your data structure. What form do we need our data to have in order to easily answer this question? (hint...`spread`)). 

## Part two: Tidying  (B)
Now let's turn our attention to the player biography `bioClean`.
  
1. Separate the `Hometown` column into two columns, namely the city and the state. Print the first few rows of these two columns. (Hint: look at the `sep=` argument of `separate`)
```{r,echo=FALSE,results='hide'}
bioClean = separate(bio, "Hometown", c("City", "State"), sep = ", " )
print(bioClean)
```
2. How many players are there from each state?
```{r}
table(bioClean$State)
```

## Pa rt three: Joining data frames

1. Find out where the offensive players are from by joining the offensive data and player biography data appropriately.

```{r}
newOff <-merge(offClean, bioClean, by = "Name")
print(newOff)
```
2. Is there a relationship between the weight of a player and the receiving yards (`Receiving_YDS`)?
```{r}
ggplot(newOff, aes(newOff$Weight, newOff$Receiving_YDS)) + xlab("Weight") + ylab("Recieving Yards") + ggtitle("Weight vs Receiving Yards") + geom_point()
```

There appears to be no coorelation between Weight and receiving yards.

3. Explore more. Raise a question that can be answered only by looking at two or more data frames. Then answer that question.
```{r}
ggplot(newOff, aes(newOff$Height, newOff$Rushing_YDS)) + xlab("Height") + ylab("Rushing Yards") + ggtitle("Weight vs Rushing Yards") + geom_point()
```
There appears to be no coorelation between height and rushing yards. 
## Tips
- Agree on a method to clean the dataset with your teammate. Both teammates can then work on the same cleaned version for downstream analysis.
- Start early to allow more time to collaborate.
- Plan ahead to prevent merge conflict: Teammates should modify the `team.Rmd` file in turn, rather than concurrently. 


