---
title: 'Template - Lab 3: Exploring Happiness'
subtitle: "Due: Thursday 10/03/2019 by 10 pm"
author: "Lynna Chu"

date: "09/26/2019"
output: html_document
---


## Background

The General Social Survey (GSS) has been run by NORC every other year since 1972 to keep track of current opinions across the United States. We are working with an excerpt of the dataset that covers years 1972 to 2016. The variable names along with a brief description are as follows: 

| Variable      | Description   |    
| ------------- |:-------------:| 
| Year          | Year for this respondent | 
| Age           | Age of respondent | 
| Sex           | Sex of respondent     | 
|....           | .....      | 

Now it's important to understand what each column stands for. You can find a codebook with explanations for each of the variables at https://gssdataexplorer.norc.org/

## Data Cleaning 

Initial exploration revealed that some data processing was needed before we could proceed with further visualization. All missing values were encoded with 'NA' as follows:
```{r}
happy <- readRDS("C:/Users/Flame/Documents/ds202/lab3/HAPPY.rds")
head(happy)
happy2 <- happy # this df has all of the missing vallues repersented with the same char
for(i in length(happy2)) happy2[i][happy2[i]=="IAP"||happy2[i]=="DK"]<- "NA"
head(happy2)
```

In total there were `....` number of missing values for the following `....` variables. 

Variables were also cast into the correct type in order to ease data visualization. This was done by:
```{r}
for(i in happy2$age) tryCatch(strtoi(happy2$age[i]))
```

Factors were also organized into meaningful order as follows: 
```{r}
head(happy2)
happySort <- happy2[order(happy2$happy,happy2$marital,happy2$degree,happy2$health,happy2$finrela,happy2$sex,happy2$age,happy2$year),]
head(happySort)
```


## Data Exploration 

Two variables of interest are `Happy` and `Health`. Their distributions are as follow: 
```{r}
library(ggplot2)
#insert relevant R code here
#include graphics. DESCRIBE your plot in context. Make sure it includes axis labels and appropriate title. 
happy <- readRDS("C:/Users/Flame/Documents/ds202/lab3/HAPPY.rds")
happy2 <- happy # this df has all of the missing vallues repersented with the same char
for(i in length(happy2)) happy2[i][happy2[i]=="IAP"||happy2[i]=="DK"]<- "NA"
ggplot(data = happy2, aes(x = happy2$HAPPY, fill = happy2$SEX) ) + geom_bar(position = position_dodge()) +xlab("Happiness") + ylab("Count of people") + ggtitle("Count of happiness of men and women") + labs(fill = "Sex")

ggplot(data = happy2, aes(x = happy2$HAPPY, fill = happy2$DEGREE) ) + geom_bar(position = position_dodge()) +xlab("Happiness") + ylab("Count of people") + ggtitle("Count of happiness by highest degree achieved") + labs(fill = "Degree Type")

ggplot(data = happy2, aes(x = happy2$HAPPY, fill = happy2$MARITAL) ) + geom_bar(position = position_dodge()) +xlab("Happiness") + ylab("Count of people") + ggtitle("Count of happiness by martial status") + labs(fill = "Marital Status")
```

The distributions for sex and happiness seem to suggest that sex does not have much of an effect on happiness. For the second graph, the graph seems to skewed left for people who are in high school or higher. For marital status, married couples are skewed left. This seems to imply that people who have finished a high school degree or higher and whose marital status is married have a better self view of their lives. 




